---
title: "The penguins of Antarctica"
author: "Selina Baldauf"
date: "`r Sys.Date()`"
format: html
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(palmerpenguins)
```

# Introduction

There are three main penguin species in Antarctica (*Chinstrap*, *Gentoo*, *Adelie*).

In this paper we want to answer the following questions

1.  How bill depth depends on bill length?
2.  Which penguin species has the highest body mass?

# Methods

## The data

The data was collected on islands in Antarctica and published by @gorman2014. You can find the original paper with the title "Ecological sexual dimorphism and environmental variability within a community of Antarctic penguins (genus *Pygoscelis*)" in PLoS ONE[^1]

[^1]: paper available [here](https://doi.org/10.1371/journal.pone.0090081).

The data is published via the `palmerpenguins` R package which you can find [on this website](https://allisonhorst.github.io/palmerpenguins/).

**The data contains (among others) the following measurements:**

-   bill length
-   bill depth
-   body mass
-   sex
    -   male
    -   female

## The analysis

We did some plots, calculated some summary statistics and a linear model of the form $y = ax + b + \epsilon$

# Results

The mean weight of all penguin species is `r mean(penguins$body_mass_g, na.rm = TRUE)` g.

The figure below shows that *Gentoo* penguins have the highest body mass.

```{r}
#| label: fig-histogram
#| fig-cap: Histogram of weight of the three penguin species.
#| fig-align: center
#| fig-width: 4
#| fig-height: 3
#| warning: false
#| message: false

ggplot(penguins, aes(x = body_mass_g, fill = species)) +
  geom_histogram(alpha = 0.6) +
  scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
  theme_minimal() +
  labs(
    x = "Body mass (g)",
    y = "Count",
    color = "Species"
  )
```

In general, it looks like the body characteristics differ between the sexes but also between the penguin species, as the table below illustrates:

```{r}
#| label: tbl-penguin
#| message: false

penguins %>%
  filter(!(is.na(sex))) %>%
  group_by(species, sex) %>%
  summarize(
    bill_length = mean(bill_length_mm, na.rm = TRUE),
    bill_depth = mean(bill_depth_mm, na.rm = TRUE),
    flipper_length = mean(flipper_length_mm, na.rm = TRUE),
    body_mass = mean(body_mass_g, na.rm = TRUE)
  )
```
