---
title: "The penguins of Antarctica"
author: "Selina Baldauf"
date: today
format: 
  html: 
    number-sections: true
    toc: true
    toc-location: left
    code-fold: true
    df-print: kable
  pdf: 
    toc: true
    number-sections: true
    df-print: kable
  docx: default
execute: 
  echo: true
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
```

# Introduction

There are three main penguin species in Antarctica (*Chinstrap*, *Gentoo*, *Adelie*).

In this paper we want to answer the following questions

1.  How bill depth depends on bill length?
2.  Which penguin species has the highest body mass?

# Methods

## The data

The data was collected on islands in Antarctica and published by Gorman et al. (2014). You can find the original paper with the title "Ecological sexual dimorphism and environmental variability within a community of Antarctic penguins (genus *Pygoscelis*)" in PLoS ONE.

The data is published via the `palmerpenguins` R package which you can find [on this website](https://allisonhorst.github.io/palmerpenguins/).

**The data contains (among others) the following measurements:**

-   bill length
-   bill depth
-   body mass
-   sex
    -   male
    -   female

## The analysis

We did some plots, calculated some summary statistics and a linear model.

# Results

The mean weight of all penguin species is `r round(mean(penguins$body_mass, na.rm = TRUE), 1)` g.

@fig-histogram below shows that *Gentoo* penguins have the highest body mass.

```{r}
#| label: fig-histogram
#| fig-cap: Histogram of weight of the three penguin species.
#| fig-align: center
#| fig-width: 4
#| fig-height: 3
#| warning: false
#| message: false

ggplot(penguins, aes(x = body_mass, fill = species)) +
  geom_histogram(alpha = 0.6) +
  scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
  theme_minimal() +
  labs(
    x = "Body mass (g)",
    y = "Count",
    fill = "Species"
  )
```

In general, it looks like the body characteristics differ between the sexes but also between the penguin species, as @tbl-penguin below illustrates:

```{r}
#| label: tbl-penguin
#| tbl-cap: Mean body characteristics of penguins
#| message: false
#| code-fold: true

penguins |>
  filter(!(is.na(sex))) |>
  summarize(
    bill_length = mean(bill_len, na.rm = TRUE),
    bill_depth = mean(bill_dep, na.rm = TRUE),
    flipper_length = mean(flipper_len, na.rm = TRUE),
    body_mass = mean(body_mass, na.rm = TRUE),
    .by = c(species, sex)
  )
```

# References